### 2020-12-19 paper/localpsf.tex
finished infinite dimensional weighting function proof

We choose the $i^\text{th}$ weighting function, $w_i(x)$, to be the solution to the following optimization problem:
\begin{equation}
\label{eq:wi_optimization_problem}
\begin{aligned}
\min_{w_i \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta w_i}_{L^2(\Omega)}^2 \\
\text{such that} &\quad w_i(x_j) = \delta_{ij}, \quad j=1,\dots,q.
\end{aligned}
\end{equation}
Here $\Delta$ is the Laplacian on $\Omega$, $\delta_{ij}$, is the Kronecker delta, which takes value one if $i=j$ and zero if $i \neq j$, and $H^2_N(\Omega)$ is the space of all functions in $H^2(\Omega)$ with Neumann zero boundary conditions. In Theorem \ref{thm:smoothest_interpolant}, we show that using these weighting functions yields the ``flattest possible'' interpolation of the convolution kernels $\varphi_i$, in a least-squares sense. In Theorem \ref{thm:wi_sum_to_one} we note that the weighting functions sum to one pointwise. These theorems rely on straightforward prerequisites that we present in Lemma \ref{lem:T_is_linear}

In Algorithm \ref{alg:weighting_functions_incremental}, we show how to efficiently solve optimization problem \eqref{eq:wi_optimization_problem}, and thus compute $w_i$. We prove correctness of Algorithm \ref{alg:weighting_functions_incremental} in Theorem \ref{thm:wi_alg_is_correct}. Constructing all $r$ of the functions $w_i$ requires solving the Poisson equation $2r$ times with different right hand side sources, and these Poisson PDE solves can be performed cheaply with multigrid. All the existing weighting functions, $\{w_i\}_{i=1}^r$, will change if a new point, $x_{r+1}$, is added. The Poisson PDE solves can be performed incrementally, so that we only need to perform two new Poisson PDE solves when a new point is added, rather than $2(r+1)$ new PDE solves.

@@

\begin{lem}
	\label{lem:T_is_linear}
	Let $f \in \mathbb{R}^r$. There exists a unique solution to the following optimization problem:
	\begin{equation}
	\label{eq:generic_optimization_initial}
	\begin{aligned}
	\min_{u \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta u}_{L^2(\Omega)}^2 \\
	\text{such that} &\quad u(x_i) = f_i, \quad i=1,\dots,r.
	\end{aligned}
	\end{equation}
	Let $\mathcal{T}:\mathbb{R}^r \rightarrow H^2(\Omega)$ denote the solution operator for optimization problem \eqref{eq:generic_optimization_initial}	as a function of $f$. That is, 
	\begin{equation*}
	\mathcal{T}(f) := u^*,
	\end{equation*}
	where $u^*$ is the minimizer of \eqref{eq:generic_optimization_initial}. The operator $\mathcal{T}$ is a linear operator.
\end{lem}

\begin{thm}
	\label{thm:smoothest_interpolant}
Let $y \in \mathbb{R}^d$, and let $u$ be the function
\begin{equation}
\label{eq:u_defn_in_thm}
u(x) := \sum_{i=1}^r \varphi_i(y)w_i(x) = \widetilde{A}\left(y+x,x\right)
\end{equation}
where the weighting functions $w_i$ are the solutions to optimization problems \eqref{eq:wi_optimization_problem}. Then $u$ solves the optimization problem
\begin{equation}
\label{eq:uy_optimization_problem}
\begin{aligned}
\min_{u \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta u}_{L^2(\Omega)}^2 \\
\text{such that} &\quad u(x_i) = \varphi_i(y), \quad i=1,\dots,q.
\end{aligned}
\end{equation}
\end{thm}
\begin{proof}
	Let $f = (\varphi_1(y), \varphi_2(y), \dots, \varphi_r(y))$ and let $e_i = (0,\dots,0,1,0,\dots,0)$ be the length $r$ vector with $i^\text{th}$ component equal to one and all other components equal to zero. The solution to optimization problem \eqref{eq:uy_optimization_problem} is given by
	\begin{align*}
		\mathcal{T}(f) &= \varphi_1(y) \mathcal{T}(e_1) + \varphi_2(y) \mathcal{T}(e_2) + \dots +\varphi_r(y) \mathcal{T}(e_r) \\
		&= \varphi_1(y) w_1 + \varphi_2(y) w_2 + \dots +\varphi_r(y) w_r \\
		&= u.
	\end{align*}
 	In the first line we used linearity of the solution map, $\mathcal{T}$ (Lemma \ref{lem:T_is_linear}), in the second line we used the fact that $w_i = \mathcal{T}(e_i)$, and in the third line we used the definition of $u$ in \eqref{eq:u_defn_in_thm}.
\end{proof}

\begin{thm}
	\label{thm:wi_sum_to_one}
	We have
	\begin{equation*}
		\sum_{i=1}^r w_i(x) = 1
	\end{equation*}
	for all $x \in \Omega$.
\end{thm}

@@

		\For{$i=1,2,\dots,r$}{
			Compute
			$$\alpha = \frac{\mathbf{1}^T S^{-1} e_i}{\mathbf{1}^T S^{-1} \mathbf{1}},$$
			where $\mathbf{1} = (1,1,\dots,1)$ is the length $r$ vector which has all components equal to one, and $e_i = (0,\dots,0,1,0,\dots,0)$ is the length $r$ vector with $i^\text{th}$ component equal to one and all other components equal to zero.

@@

\begin{thm}
\label{thm:wi_alg_is_correct}
	The weighting functions $w_i$ computed by Algorithm \ref{alg:weighting_functions_incremental} solve optimization problem \ref{eq:wi_optimization_problem}.
\end{thm}

@@

	In this proof, we will repeatedly make use of the fact that
	\begin{equation}
	\label{eq:inline_wi_lemma}
	\left(\psi_j, \Delta h\right)_{L^2(\Omega)} = h(x_j)
	\end{equation}
	for all $h\in H^2_N(\Omega)$. This follows from integration by parts and the definition of $\psi$ in \eqref{eq:psi_eq}.

	From standard optimization theory, the solution to optimization problem \ref{eq:wi_optimization_problem} is the stationary point of the Lagrangian,
	\begin{equation*}
		\mathcal{L}(w,\lambda) = \frac{1}{2}\|\Delta w\|^2 + \sum_{j=1}^r \lambda_j\left(e_j - w(x_j)\right).
	\end{equation*}
	Using the definition of $w_i$ in \eqref{eq:definition_of_wi_alg}, we may write the gradient of $\mathcal{L}$ with respect to $\lambda_j$ as
	\begin{equation*}
		\nabla_{\lambda_j} \mathcal{L}(w, \lambda) = e_j - w(x_j) 
		= e_j - \alpha + \sum_{k=1}^r \lambda_k \theta_k(x_j).
	\end{equation*}
	Using \eqref{eq:inline_wi_lemma} with $h=\theta_k$, the equation for $\theta$ in \eqref{eq:theta_eq}, and the definition of $S$ in \eqref{eq:defn_of_S_alg}, we have
	\begin{align*}
		\sum_{k=1}^r \lambda_k \theta_k(x_j) &= \sum_{k=1}^r \lambda_k \left(\psi_j, \Delta \theta_k\right)_{L^2(\Omega)} \\
		&= \sum_{k=1}^r \lambda_k \left(\psi_j, \psi_k\right)_{L^2(\Omega)} = \left(S \lambda\right)_j.
	\end{align*}
	Combining all of these components of $\nabla_\lambda \mathcal{L}$ into a single vector and using the definition of $\lambda$ in \eqref{eq:definition_of_mu_alg} yields
	\begin{equation*}
		\nabla_\lambda \mathcal{L}(w, \lambda) = e - \alpha \mathbf{1} + S \lambda = e - \alpha \mathbf{1} + S \left(-S^{-1}\left(e - \alpha \mathbf{1}\right)\right) = 0.
	\end{equation*}

	Taking the directional derivative of $\mathcal{L}$ with respect to $w$ in direction $h$ and using \eqref{eq:inline_wi_lemma} yields
	\begin{align*}
		\nabla_w \mathcal{L}(w, \lambda)h &= \left(\Delta w, \Delta h\right)_{L^2(\Omega)} - \sum_{j=1}^r \lambda_j h(x_j) \\
		&= \left(\Delta w, \Delta h\right)_{L^2(\Omega)} - \sum_{j=1}^r \lambda_j \left(\psi_j, \Delta h\right)_{L^2(\Omega)} \\
		&= \left(\Delta w - \sum_{j=1}^r \lambda_j \psi_j,~ \Delta h\right)_{L^2(\Omega)}.
	\end{align*}
	Using the definition of $w$ in Algorithm \ref{alg:weighting_functions_incremental}, and the fact that the Laplacian of a constant function is zero, we have
	\begin{align*}
		\Delta w &= -\sum_{j=1}^r \lambda_j \Delta \theta_j \\
		&= -\sum_{j=1}^r \lambda_j \psi_j
	\end{align*}

>>>>

We choose the $i^\text{th}$ weighting function, $w_i(x)$, to be the solution to the following optimization problem:
\begin{equation}
\label{eq:wi_optimization_problem}
\begin{aligned}
\min_{w_i \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta w_i}_{L^2(\Omega)}^2 \\
\text{such that} &\quad w_i(x_j) = \delta_{ij}, \quad j=1,\dots,q.
\end{aligned}
\end{equation}
Here $\Delta$ is the Laplacian on $\Omega$, $\delta_{ij}$, is the Kronecker delta, which takes value one if $i=j$ and zero if $i \neq j$, and $H^2_N(\Omega)$ is the space of all functions in $H^2(\Omega)$ with Neumann zero boundary conditions. In Theorem \ref{thm:smoothest_interpolant}, we show that using these weighting functions yields the ``flattest possible'' interpolation of the convolution kernels $\varphi_i$, in a least-squares sense. In Proposition \ref{thm:wi_sum_to_one} we note that the weighting functions sum to one pointwise. These theorems rely on straightforward prerequisites that we present in Lemma \ref{lem:T_is_linear}

In Algorithm \ref{alg:weighting_functions_incremental}, we show how to efficiently solve optimization problem \eqref{eq:wi_optimization_problem}, and thus compute $w_i$. We prove correctness of Algorithm \ref{alg:weighting_functions_incremental} in Theorem \ref{thm:wi_alg_is_correct}. Theorem \ref{thm:wi_alg_is_correct} relies on a prerequisite result which we present in Lemma \ref{lem:psi_delta_eval}. Constructing all $r$ of the functions $w_i$ requires solving the Poisson equation $2r$ times with different right hand side sources, and these Poisson PDE solves can be performed cheaply with multigrid. All the existing weighting functions, $\{w_i\}_{i=1}^r$, will change if a new point, $x_{r+1}$, is added. The Poisson PDE solves can be performed incrementally, so that we only need to perform two new Poisson PDE solves when a new point is added, rather than $2(r+1)$ new PDE solves.

@@

\begin{lem}[Preparation for Theorem \ref{thm:smoothest_interpolant}]
	\label{lem:T_is_linear}
	Let $f \in \mathbb{R}^r$. There exists a unique solution to the following optimization problem:
	\begin{equation}
	\label{eq:generic_optimization_initial}
	\begin{aligned}
	\min_{u \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta u}_{L^2(\Omega)}^2 \\
	\text{such that} &\quad u(x_i) = f_i, \quad i=1,\dots,r.
	\end{aligned}
	\end{equation}
	Let $\mathcal{T}:\mathbb{R}^r \rightarrow H^2(\Omega)$ denote the solution operator for optimization problem \eqref{eq:generic_optimization_initial}	as a function of $f$. That is, 
	\begin{equation*}
	\mathcal{T}(f) := u^*,
	\end{equation*}
	where $u^*$ is the minimizer of \eqref{eq:generic_optimization_initial}. The operator $\mathcal{T}$ is a linear operator.
\end{lem}

\begin{thm}[Optimal interpolation]
	\label{thm:smoothest_interpolant}
Let $y \in \mathbb{R}^d$, and let $u$ be the function
\begin{equation}
\label{eq:u_defn_in_thm}
u(x) := \sum_{i=1}^r \varphi_i(y)w_i(x) = \widetilde{A}\left(y+x,x\right)
\end{equation}
where the weighting functions $w_i$ are the solutions to optimization problems \eqref{eq:wi_optimization_problem}. Then $u$ solves the optimization problem
\begin{equation}
\label{eq:uy_optimization_problem}
\begin{aligned}
\min_{u \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta u}_{L^2(\Omega)}^2 \\
\text{such that} &\quad u(x_i) = \varphi_i(y), \quad i=1,\dots,q.
\end{aligned}
\end{equation}
\end{thm}
\begin{proof}
	Let $f = (\varphi_1(y), \varphi_2(y), \dots, \varphi_r(y))$ and let $e_i = (0,\dots,0,1,0,\dots,0)$ be the length $r$ vector with $i^\text{th}$ component equal to one and all other components equal to zero. The solution to optimization problem \eqref{eq:uy_optimization_problem} is given by
	\begin{align*}
		\mathcal{T}(f) &= \varphi_1(y) \mathcal{T}(e_1) + \varphi_2(y) \mathcal{T}(e_2) + \dots +\varphi_r(y) \mathcal{T}(e_r) \\
		&= \varphi_1(y) w_1 + \varphi_2(y) w_2 + \dots +\varphi_r(y) w_r \\
		&= u.
	\end{align*}
 	In the first line we used linearity of the solution map, $\mathcal{T}$ (Lemma \ref{lem:T_is_linear}), in the second line we used the fact that $w_i = \mathcal{T}(e_i)$, and in the third line we used the definition of $u$ in \eqref{eq:u_defn_in_thm}.
\end{proof}

\begin{prop}
	\label{thm:wi_sum_to_one}
	We have
	\begin{equation*}
		\sum_{i=1}^r w_i(x) = 1
	\end{equation*}
	for all $x \in \Omega$.
\end{prop}

@@

		\For{$i=1,2,\dots,r$}{
			Compute
			\begin{equation}
			\label{eq:definition_of_alpha_alg}
			\alpha = \frac{\mathbf{1}^T S^{-1} e_i}{\mathbf{1}^T S^{-1} \mathbf{1}},
			\end{equation}
			where $\mathbf{1} = (1,1,\dots,1)$ is the length $r$ vector which has all components equal to one, and $e_i = (0,\dots,0,1,0,\dots,0)$ is the length $r$ vector with $i^\text{th}$ component equal to one and all other components equal to zero.

@@

\begin{lem}[Preparation for Theorem \ref{thm:wi_alg_is_correct}]
	\label{lem:psi_delta_eval}
	We have
	\begin{equation}
	\label{eq:inline_wi_lemma}
	\left(\psi_j, \Delta h\right)_{L^2(\Omega)} = h(x_j) - \avg(h)
	\end{equation}
	for all $h\in H^2_N(\Omega)$, where
	\begin{equation*}
		\avg(h) :=	 \frac{1}{|\Omega|} \int_\Omega h ~dx.
	\end{equation*}
\end{lem}

\begin{proof}
	This follows from integration by parts and the definition of $\psi$ in \eqref{eq:psi_eq}.
\end{proof}

\begin{thm}[Correctness of Algorithm \ref{alg:weighting_functions_incremental}]
\label{thm:wi_alg_is_correct}
	The weighting functions $w_i$ computed by Algorithm \ref{alg:weighting_functions_incremental} solve optimization problem \ref{eq:wi_optimization_problem}.
\end{thm}

@@

	From standard optimization theory, the solution to optimization problem \ref{eq:wi_optimization_problem} is the stationary point of the Lagrangian,
	\begin{equation*}
		\mathcal{L}(w,\lambda) = \frac{1}{2}\|\Delta w\|^2 + \sum_{j=1}^r \lambda_j\left(w(x_j) - e_j\right).
	\end{equation*}
	Using the definition of $w_i$ in \eqref{eq:definition_of_wi_alg}, we may write the gradient of $\mathcal{L}$ with respect to $\lambda_j$ as
	\begin{equation*}
		\nabla_{\lambda_j} \mathcal{L}(w, \lambda) = w(x_j) - e_j
		= \alpha - \sum_{k=1}^r \lambda_k \theta_k(x_j) - e_j.
	\end{equation*}
	Using Lemma \ref{lem:psi_delta_eval} with $h=\theta_k$, the equation for $\theta_k$ in \eqref{eq:theta_eq}, and the definition of $S$ in \eqref{eq:defn_of_S_alg}, we have
	\begin{align*}
		\sum_{k=1}^r \lambda_k \theta_k(x_j) &= \sum_{k=1}^r \lambda_k \left(\psi_j, \Delta \theta_k\right)_{L^2(\Omega)}\\
		&= \sum_{k=1}^r \lambda_k \left(\psi_j, \psi_k\right)_{L^2(\Omega)} = \left(S \lambda\right)_j.
	\end{align*}
	Combining all of these components of $\nabla_\lambda \mathcal{L}$ into a single vector and using the definition of $\lambda$ in \eqref{eq:definition_of_mu_alg} yields
	\begin{align*}
		\nabla_\lambda \mathcal{L}(w, \lambda) &= \alpha \mathbf{1} - e - S \lambda \\
		&= \alpha \mathbf{1} - e - S \left(-S^{-1}\left(e - \alpha \mathbf{1}\right)\right) = 0.
	\end{align*}

	Taking the directional derivative of $\mathcal{L}$ with respect to $w$ in direction $h$ and using Lemma \ref{lem:psi_delta_eval} yields
	\begin{align}
		\nabla_w \mathcal{L}(w, \lambda)h &= \left(\Delta w, \Delta h\right)_{L^2(\Omega)} + \sum_{j=1}^r \lambda_j h(x_j) \nonumber \\
		&= \left(\Delta w, \Delta h\right)_{L^2(\Omega)} + \sum_{j=1}^r \lambda_j \left(\psi_j, \Delta h\right)_{L^2(\Omega)} + \lambda_j \avg(h) \nonumber \\
		&= \left(\Delta w + \sum_{j=1}^r \lambda_j \psi_j,~ \Delta h\right)_{L^2(\Omega)} + \avg(h) \mathbf{1}^T \lambda. \label{eq:L_w_38}
	\end{align}
	Using the definition of $w$ in Algorithm \ref{alg:weighting_functions_incremental}, and the fact that the Laplacian of a constant function is zero, we have
	\begin{equation*}
		\Delta w = -\sum_{j=1}^r \lambda_j \Delta \theta_j 
		= -\sum_{j=1}^r \lambda_j \psi_j,
	\end{equation*}
	which implies that the first term in \eqref{eq:L_w_38} is zero. By the definition of $\lambda$ in \eqref{eq:definition_of_mu_alg} and $\alpha$ in \eqref{eq:definition_of_alpha_alg}, we have
	\begin{align*}
		\mathbf{1}^T \lambda &= -\mathbf{1}^T S^{-1} \left(e - \alpha \mathbf{1}\right) \\
		&= -\mathbf{1}^T S^{-1} e + \alpha \mathbf{1}^T S^{-1} \mathbf{1} \\
		&= -\mathbf{1}^T S^{-1} e + \frac{\mathbf{1}^T S^{-1} e}{\mathbf{1}^T S^{-1} \mathbf{1}} \mathbf{1}^T S^{-1} \mathbf{1} = 0,
	\end{align*}
	which implies that the second term in \eqref{eq:L_w_38} is zero. Thus $\nabla_w \mathcal{L}(w, \lambda) = 0$. Since $\nabla \mathcal{L}(w, \lambda)=0$, Algorithm \ref{alg:weighting_functions_incremental} correctly produces the solution to optimization problem \ref{eq:wi_optimization_problem}.
\end{proof}
